Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-07-11 manuscript/manuscript.tex
Started writing Theory

It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ degree. 
Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.  \citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
(1) a faster calculation of the distance from the computation point to the tesseroid and 
(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.

@@

>>>>

It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ order. 
Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.  \citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
(1) a faster calculation of the distance from the computation point to the tesseroid and 
(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.

@@

The spherical prisms known as tesseroids are mass elements defined in a geocentric spherical coordinate system bounded by a pair of parallels, a pair of  meridians, and two concentric spherical surfaces.
We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented cartesian coordinate system located at $P$.
In the special case of homogeneous density, the gravity potential, gradient and Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.

In our case, we will suppose that the tesseroid has an arbitrary variable density in depth, this means that only depends on the radius spherical coordinate. Thus, the integrals for the gravity fields are slightly modified.

\begin{equation}
    V(r,\phi,\lambda) = G
    \int\limits_{\lambda_1}^{\lambda_2}
    \int\limits_{\phi_1}^{\phi_2}
    \int\limits_{r_1}^{r_2}
    \frac{\rho(r')}{\ell} \kappa \,  dr' d\phi' d\lambda',
\label{eq:tesspot}
\end{equation}
\begin{equation}
    g_{\alpha}(r,\phi,\lambda) = G
    \int\limits_{\lambda_1}^{\lambda_2}
    \int\limits_{\phi_1}^{\phi_2}
    \int\limits_{r_1}^{r_2}
    \rho(r') \frac{\Delta_\alpha}{\ell^3}
    \kappa \, dr' d\phi' d\lambda',
\label{eq:tessgrav}
\end{equation}
\begin{equation}
    g_{\alpha\beta}(r,\phi,\lambda) = G
    \int\limits_{\lambda_1}^{\lambda_2}
    \int\limits_{\phi_1}^{\phi_2}
    \int\limits_{r_1}^{r_2}
    \rho(r') I_{\alpha\beta} \, \kappa \, dr' d\phi' d\lambda' ,
    \label{eq:tesstensor}
\end{equation}
\noindent where
\begin{equation}
    I_{\alpha\beta} =
    \left(
        \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
        \frac{\delta_{\alpha\beta}}{\ell^3}
    \right) ,
    \label{eq:tesstensorkernel}
\end{equation}

\noindent $\alpha, \beta \in \{x, y, z\}$,
$\rho(r')$ is the density function that depends on the radius coordinate,
$\delta_{\alpha\beta}$ is Kronecker's delta,
$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
\begin{equation}
    \Delta_x = r'[\cos\phi\sin\phi' - \sin\phi\cos\phi'
               \cos(\lambda' - \lambda)],
\end{equation}
\begin{equation}
    \Delta_y = r' \cos \phi' \sin(\lambda' - \lambda),
\end{equation}
\begin{equation}
    \Delta_z = r' \cos \psi - r,
\end{equation}
\begin{equation}
    \kappa = {r'}^2 \cos \phi',
\end{equation}
\begin{equation}
    \ell = \sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi},
\end{equation}
\begin{equation}
    \cos\psi = \sin\phi\sin\phi' + \cos\phi\cos\phi'
                 \cos(\lambda' - \lambda).
\end{equation}

According to \citet[p.390]{Hildebrand1987}, we can approximate any integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e. by a weighted sum of the integration kernel evaluated on the roots of the $N$th Legendre polynomial.

\begin{equation}
    \int\limits_{-1}^1 f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i),
\end{equation}

\noindent where $x_i$ are the roots of the $N$th Legendre polynomial $P_N(x)$ and $W_i$ are the weights of each term.

\begin{equation}
    W_i = \frac{2}{(1-x_i^2)[P_N^\prime(x_i)]^2},
\end{equation}

\begin{equation}
    P_N(x_i) = 0, \quad \forall i = {1,\dots,N}.
\end{equation}

In case our integration is defined on a different interval, e.g. $[a,b]$, we can scale the nodes and perform a similar approximation.

\begin{equation}
    \int\limits_a^b f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i^s),
\end{equation}

\noindent where $x_i^s$ is the scaled $x_i$ root in the $[a,b]$ interval, also called nodes of the quadrature:

\begin{equation}
    x_i^s = \frac{b-a}{2} x_i + \frac{b+a}{2}.
\end{equation}

We are going to use only a second order GLQ, so the roots of the $P_2(x)$ Legendre polynomial are $x_i = \pm 0.577350269$ and the weights are $W_i = 1$.

Our intention is to use GLQ to approximate the gravity fields volume integrals from equations \ref{eq:tesspot}-\ref{eq:tesstensor}. Applying the GLQ definition for every integral corresponding to each spherical coordinate we obtain the following expression \citep{Asgharzadeh2007, Uieda2016}:

\begin{equation}
    \iiint\limits_\Omega f(r', \phi', \lambda') d\Omega \approx A 
    \sum\limits_{i=1}^{N^r}
    \sum\limits_{j=1}^{N^\phi}
    \sum\limits_{k=1}^{N^\lambda}
    W_i^r W_j^\phi W_k^\lambda f(r_i, \phi_j, \lambda_k),
\end{equation}

\noindent where

\begin{equation}
    A = 
    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}.
\end{equation}

In our case with variable density, we can write the integral kernel as:

\begin{equation}
    f(r', \phi', \lambda') = \rho(r') g(r', \phi', \lambda')
\end{equation}
